Linear instability underlying synchronous waving of marine grass in flow

### 2014-09-01 Paper/grass.tex
grass bbl

\begin{figure}
\begin{center}
\includegraphics[]{Critical_Re_vs_delta_noshear} \\
\vspace{-6mm} \hspace{-3mm}
\includegraphics[]{K_vs_shear_width_noshear}
\end{center}
\caption{Critical Reynolds number with flow (top) and the corresponding marginally stable wave number (bottom) for different submergence ration as a function of vegetation density
 parametrized by the boundary layer thickness. Parameters estimated from experiments reported by Ghisalberti and Nepf\cite{Ghisal02} to exhibit or suppress synchronous waving are 
also included in the top panel. In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed. Inset compares experimental observations 
of the dominant frequency $f_o$ (in Hz) measured with the predictions from the solution of \eqref{Orr-somerfield}. The experimental data in the inset in obtained from the aricles by 
Ghisalberti \cite{Ghisal02} and Vivoni \cite{Vivoni98}.}
\label{Re_vs_delta}
\end{figure}
We use a mean field model for the coupling between the flow and the canopy. 
Drag force exerted by the grass is approximated via a continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance\cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06} as
\begin{equation}
\rho \left(\bu_{t}+\bu.\grad\bu \right) = -\grad P+\mu\grad^{2}\bu +\mathbf{f}+\rho\mathbf{g}, \quad \grad\cdot\bu = 0,
\label{navier-stokes}
\end{equation}
where $\mathbf{f_{d}}$ is drag force per unit length of the grass blade, $N_g$ the grass number density per unit area, $\rho$ the fluid density, $\mathbf{u}$ the velocity, 
$P$ the pressure, $\mu$ the turbulent viscosity and $\mathbf{g}$ the acceleration due to gravity. 
The Reynolds number of the flow on the scale of a grass blade is expected to be O($10^2$-$10^3$), and hence the drag force is dominated by form drag, and is modeled as 
$\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\bn + C_{T}\rho u_{T}^{2}d\bt$ \shreyas{citation needed} 
where $d$ is average width of grass blade $C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively; $\bu_{T}$, $\bu_{N}$ are velocity vector along and
normal to grass while $\bt,\bn$ being unit vector along and normal to grass. 
We expect $C_T \ll C_N$ and take $C_T=0$ for rest of analysis. 
In the field, both $C_N$ and $N_g$ vary with distance from the bottom but we do not expect these variation to be central to the mechanism and therefore take $C_N$ and $N_g$ to be constants. 
Based on previous experiments\cite{Vivoni98}, we take $C_N \approx 1$.

@@

The distinct asymptotic behavior of the two modes as $\Ndg \gg 1$ allows us to understand the mechanism of the instability. 
Mode 2 has a threshold $\Rey \sim ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in figure \ref{Re_vs_delta}, which can be understood by asymptotic analysis of \eqref{Orr-somerfield}.
Noting that for large $\Ndg$, the non-dimensional steady flow velocity inside the grass is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
Using this estimate, and $\Rey\gg1$, \eqref{Orr-somerfield} within the grass simplifies into $\sigma\left(D^2-k^2\right)\phi = -2{(\Ndg/\Rey)^{1/2}}D^2\phi$ \shreyas{check sign}, while in the unvegetated region it simplifies to the Rayleigh equation $ \left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi$. 
This analysis predicts that the mode shape would converge as $\Rey\gg 1$ and $\Ndg \gg 1$ but $\Rey/\Ndg \sim O(1)$, in agreement with out numerical results shown in figure \ref{Asymptotic_mode}. 
We interpret this as the instability of an inviscid flow, with the vegetation modeled by a continuum drag field, and for which the boundary layer near the grass tip plays no role. The only remaining parameter $\Rey/\Ndg$ sets the threshold, leading to the asymptotic behavior $\Rey \propto \Ndg$ (or $\Rey \sim ({\delta}/{H})^{-3/2}$).

\begin{figure}
\includegraphics[]{Asymptotic_noshear}
\caption{plot of $|\phi|$ for different shear layer thickness in limit of small shear layer thickness}
\label{Asymptotic_mode}
\end{figure}

>>>>

\begin{figure}
\begin{center}
\includegraphics[]{Critical_Re_vs_delta_noshear} \\
\vspace{-6mm} \hspace{-3mm}
\includegraphics[]{K_vs_shear_width_noshear}
\end{center}
\caption{Critical Reynolds number with flow (top) and the corresponding marginally stable wave number (bottom) for different submergence ratio as a function of vegetation density
 parametrized by the boundary layer thickness. Parameters estimated from experiments reported by Ghisalberti and Nepf\cite{Ghisal02} to exhibit or suppress synchronous waving are 
also included in the top panel. In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed. Inset compares experimental observations 
of the dominant frequency $f_o$ (in Hz) measured with the predictions from the solution of \eqref{Orr-somerfield}. The experimental data in the inset in obtained from the articles by 
Ghisalberti \cite{Ghisal02} and Vivoni \cite{Vivoni98}.}
\label{Re_vs_delta}
\end{figure}
We use a mean field model for the coupling between the flow and the canopy. 
Drag force exerted by the grass is approximated via a continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance\cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06} as
\begin{equation}
\rho \left(\bu_{t}+\bu.\grad\bu \right) = -\grad P+\mu\grad^{2}\bu +\mathbf{f}+\rho\mathbf{g}, \quad \grad\cdot\bu = 0,
\label{navier-stokes}
\end{equation}
where $\mathbf{f_{d}}$ is drag force per unit length of the grass blade, $N_g$ the grass number density per unit area, $\rho$ the fluid density, $\mathbf{u}$ the velocity, 
$P$ the pressure, $\mu$ the turbulent viscosity and $\mathbf{g}$ the acceleration due to gravity. 
The Reynolds number of the flow on the scale of a grass blade is expected to be O($10^2$-$10^3$), and hence the drag force is dominated by form drag, and is modeled as 
$\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\bn + C_{T}\rho u_{T}^{2}d\bt$ \shreyas{citation needed} 
where $d$ is average width of grass blade $C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively; $\bu_{T}$, $\bu_{N}$ are velocity vector along and
normal to grass while $\bt,\bn$ being unit vector along and normal to grass. 
We expect $C_T \ll C_N$ and take $C_T=0$ for rest of analysis. 
In the field, both $C_N$ and $N_g$ vary with distance from the bottom but we do not expect these variation to be central to the mechanism and therefore take $C_N$ and $N_g$ to be constants. 
Based on previous experiments\cite{Vivoni98}, we take $C_N \approx 1$.

@@

The distinct asymptotic behavior of the two modes as $\Ndg \gg 1$ allows us to understand the mechanism of the instability. 
Mode 2 has a threshold $\Rey \sim ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in figure \ref{Re_vs_delta}, which can be understood by asymptotic analysis of \eqref{Orr-somerfield}.
Noting that for large $\Ndg$, the non-dimensional steady flow velocity inside the grass is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
Using this estimate, and $\Rey\gg1$, \eqref{Orr-somerfield} within the grass simplifies into $\sigma\left(D^2-k^2\right)\phi = -2{(\Ndg/\Rey)^{1/2}}D^2\phi$ \shreyas{check sign}, while in the unvegetated region it simplifies to the Rayleigh equation $ \left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi$. 
This analysis predicts that the mode shape would converge as $\Rey\gg 1$ and $\Ndg \gg 1$ but $\Rey/\Ndg \sim O(1)$, in agreement with out numerical results shown in figure \ref{Asymptotic_mode}.

We interpret this as the instability of an inviscid flow, with the vegetation modeled by a continuum drag field, and for which the boundary layer near the grass tip plays no role. The only remaining parameter $\Rey/\Ndg$ sets the threshold, leading to the asymptotic behavior $\Rey \propto \Ndg$ (or $\Rey \sim ({\delta}/{H})^{-3/2}$).

\begin{figure}
\includegraphics[]{Asymptotic_noshear}
\caption{Plot of mode shape $|\phi|$ for different shear layer thickness in limit of small shear layer thickness for representative values of grass density. Mode 1 is shown in solid and Mode 2 is shown in dashed. The parameters Mode 2 shapes are chosen such that $\Rey \gg 1$, $\Ndg \gg 1$ (specifying in terms of $\delta/H$) but $\Rey/\Ndg = O(1)$. The mode shapes approach each other for these small values of $\delta/H$ indicating that the large grass density asymptote is reached. Mode 2 shapes appear self-similar in shape as $\delta\to 0$, and are compared to each other on a rescaled coordinate in the inset. Inset shows $|\phi|$ for Mode 1 as a function of $(y-\hg)/\delta$. The modes approach a universal shape, indicating that an asymptotic limit has been reached. The limit is not yet reached for the case $\delta/H = 0.01$ due to the influence of bottom boundary; the grass height in this case is comparable to the boundary layer thickness.}
\label{Asymptotic_mode}
\end{figure}
